---
title: "Lesson3-Univariate Problems"
output: github_document
---

##### Keywords
file format, histogram, facet, scaling, frequency polygon

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Import libraries and read Dimond data set
```{r message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(tidyr)
library(dplyr)

data("diamonds")
```

### Problem 1. number of observation/variables, data types
```{r}
summary(diamonds)
```

```{r}
names(diamonds)
```

```{r}
dim(diamonds)
```

```{r}
head(diamonds)
```

```{r}
levels(diamonds$color)
```

### Problem 2. histogram
```{r}
qplot(x = price, data = diamonds,
      binwidth = 200,
      color = I('black'), fill = I('blue'),
      xlab = "Price Range",
      ylab = "Number of Diamonds") +
  scale_x_continuous(breaks = seq(0, 100, 20000))
```

### Problem 3. histogram summary
```{r}
summary(diamonds$price)
```
From the histogram, it's long-tailed having median on the left of mean. The summary of the price supports the fact that most of the prices are centered around 0 ~ 2500.

### Problem 4. how many diamonds in a certain range
```{r}
diamonds$price_range <- ifelse(diamonds$price < 250, 'cheap', 'average')
diamonds$price_range <- ifelse(diamonds$price < 500, diamonds$price_range, 'expensive')
diamonds$price_range <- ifelse(diamonds$price < 15000, diamonds$price_range, 'premium')

table(diamonds$price_range)
```

### Problem 5. peak of histogram
```{r}
# round the prices (e.g. 133 -> 100, 271 -> 300)
price_rounded <- round(diamonds$price, digits = -2)
# list prices and find the most found
uniq_price <- unique(price_rounded)
uniq_price[which.max(tabulate(match(price_rounded, uniq_price)))]
# count the most found
length(price_rounded[price_rounded == 700])
```

### Problem 6. price by cut histogram
```{r}
p1 <- qplot(x = price, data = subset(diamonds, cut == 'Fair'), xlab = 'Fair')
p2 <- qplot(x = price, data = subset(diamonds, cut == 'Good'), xlab = 'Good')
p3 <- qplot(x = price, data = subset(diamonds, cut == 'Very Good'), xlab = 'Very Good')
p4 <- qplot(x = price, data = subset(diamonds, cut == 'Premium'), xlab = 'Premium')
p5 <- qplot(x = price, data = subset(diamonds, cut == 'Ideal'), xlab = 'Ideal')

grid.arrange(p1, p2, p3, p4, p5)
```

### Problem 7. Price by cut
```{r}
by(diamonds$price, diamonds$cut, summary)
```

### Problem 8. scale and multiple histogram
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(color = 'black', fill = 'red') +
  scale_x_continuous(limits = c(0, 20000), breaks = seq(0, 20000, 1000)) +
  theme_bw(base_size = 5) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~cut, scales = 'free')
```

### Problem 9. price per Carat by cut
```{r warning=FALSE}
ggplot(aes(x = price/carat), data = diamonds) +
  geom_histogram(color = 'black', fill = 'blue') +
  scale_x_log10(limits = c(500, 15000), 
                breaks = scales::trans_breaks('log10', function(x) 10^x)) +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  facet_wrap(~cut, scales = 'free')
```

### Problem 10. price box plot
```{r}
# price by clarity
qplot(x = clarity, y = price,
      data = diamonds,
      geom = 'boxplot')
```

```{r}
# price by cut
qplot(x = cut, y = price,
      data = diamonds,
      geom = 'boxplot')
```

```{r}
# price by color
ggplot(aes(x = color, y = price), data = diamonds) +
  geom_boxplot()
```

### Problem 11. IQR
```{r}
# price by color summary
IQR(subset(diamonds, color == 'D')$price)
```

### Problem 12. price per carat using box plot
```{r}
qplot(x = color, y = price/carat,
      data = diamonds,
      geom = 'boxplot') +
  coord_cartesian(ylim = c(2000, 6000))
```

The best color is 'D' and the next to the right and so on, to the worst 'J' on the x-axis. Surprisingly enough, the best color seems to be less than average in median price and the worst color is one of the high median prices. And 'G' in the middle has the highest price. All these imply that the factor color is not actually influential on price of a diamond.

### Problem 13. crat frequency polygon
```{r}
ggplot(aes(x = carat), data = diamonds) +
  geom_freqpoly(binwidth = 0.01)
```

```{r}
table(diamonds$carat == 1.01)
```