---
title: "Mini-Project Series I: The Trends of GDP/Capita By Countries"
output: github_document
---

### Introduction
The goal of this project series is to explore some macro economic data mostly by visualizing for analysis. For that, there will be a lot of data processing such as creating new variables as well as handling incomplete information. The first topic of the series is about GDP/capita which is one of the well-known economic measures and is what I personally look at when surveying a country. The data set from [Gap Minder](https://www.gapminder.org/data/) is a matrix with inflation-adjusted GDP/capita from 1960 to 2011 for plenty of countries. Histograms of the past and the present will be drawn to see how the countries have grown in the long term. For the recent trends, I will calculate the growth rate in recent 5 years for each nation to see at what rate most countries have grown and which country grew the most and the least. Lastly, the countries will be classified into 3 groups by GDP/capita and will be visualized to show any trend they have.

### Import libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
library(xlsx)
library(ggplot2)
library(tidyr)
library(dplyr)
library(gridExtra)
```

### Data from Gap Minder: [GDP/capita (US$, inflation-adjusted) by World Bank](https://docs.google.com/spreadsheets/d/1RctTQmKB0hzbm1E8rGcufYdMshRdhmYdeL29nXqmvsc/pub)
```{r}
gdp_capita <- read.csv('../Data/GDPpercapitaconstant2000US.csv')
colnames(gdp_capita)[1] <- "Country"
```

Even if the data set starts from 1960, there are too many missing values to be a quality source. For this reason, the values from 1970 will be used for analysis.

### Histogram of The Past Year (1970) vs. The Recent Year (2011)
```{r warning=FALSE}
p_1970 <- ggplot(aes(x = X1970), data = subset(gdp_capita, !is.na(X1960))) +
  geom_histogram(binwidth = 500, color = 'gray', fill = 'lightblue') +
  xlab("GDP/capita in 1970") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(limits = c(0, 55000), breaks = seq(0, 55000, 2000))

p_2011 <- ggplot(aes(x = X2011), data = subset(gdp_capita, !is.na(X1960))) +
  geom_histogram(binwidth = 500, color = 'gray', fill = 'lightblue') +
  xlab("GDP/capita in 2011") +
  theme_bw(base_size = 10) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_continuous(limits = c(0, 55000), breaks = seq(0, 55000, 2000))

grid.arrange(p_1970, p_2011, ncol = 1)
```

Unlike in 1970, there are many countries with GDP/capita over 20,000 USD in 2011. However, there are still many remaining under 2,000 USD. In conclusion, it can be said that the GDP/capita has moved up in the last a few decades but the disparity among the countries must have also grown.

```{r}
gdp_capita$Country[gdp_capita$X2011 > 50000 & !is.na(gdp_capita$X2011)]
```

### Countries by Growth
```{r warning=FALSE}
# calculate 2006~2011 growth rate
gdp_capita_5yr <- subset(gdp_capita, !is.na(gdp_capita$X2006) & !is.na(gdp_capita$X2011))
gdp_capita_5yr <- select(gdp_capita_5yr, matches("Country"), X2006:X2011)
gdp_capita_5yr$growth_rate <- ((gdp_capita_5yr$X2011 - gdp_capita_5yr$X2006) 
                               / gdp_capita_5yr$X2006) * 100

gdp_capita_5yr_hist <- ggplot(aes(x = growth_rate), data = gdp_capita_5yr) +
  geom_histogram(binwidth = 1, color = 'gray', fill = 'lightgreen') +
  xlab("5 Years Growth Rate") + ylab("Countries")
  
gdp_capita_5yr_hist + scale_x_continuous(limits = c(-30, 70), breaks = seq(-30, 70, 5))
```

The 5 years growth rate is how much each GDP/capita has grown from 2006 to 2011. As shown, most of the countries fall between -10% and 10%. The reason why there are many with minus growth is supposedly because of the global financial crisis of 2007~2008. With an oil-producing country of UAE ranked the least growth, many EU countries are listed in the minus growth countries, too. On the other hand, the countries with more than 20% growth in the same period are mostly undeveloped countries in Africa and Asia.

```{r}
summary(gdp_capita_5yr$growth_rate)
```

```{r}
gdp_capita_5yr$Country[gdp_capita_5yr$growth_rate == min(gdp_capita_5yr$growth_rate)]
```

[UAE GDP/capita without inflation adjustment](https://www.google.com/search?rlz=1C1CAFA_enKR676KR676&ei=muUdW5_BMtGOsQXaiKTQBA&q=UAE+gdp+per+capita&oq=UAE+gdp+per+capita&gs_l=psy-ab.3..0i67k1j0i7i30k1j0l2j0i30k1l4j0i10i30k1j0i30k1.21108.21587.0.21827.3.3.0.0.0.0.105.279.2j1.3.0....0...1c.1.64.psy-ab..0.3.279...0i7i10i30k1j0i10k1.0.ZGpxk3626Nc)

```{r}
gdp_capita_5yr$Country[gdp_capita_5yr$growth_rate == max(gdp_capita_5yr$growth_rate)]
```

[Tajikistan GDP/capita without inflation adjustment](https://www.google.com/search?rlz=1C1CAFA_enKR676KR676&ei=eOUdW6CvNIyWsgW08424CQ&q=tajikistan+gdp+per+capita&oq=tajikis&gs_l=psy-ab.3.0.0i67k1l8j0j0i67k1.28516.30091.0.33115.7.6.0.1.1.0.100.525.5j1.6.0....0...1c.1.64.psy-ab..0.7.528...35i39k1j0i131i67k1j0i131k1j0i20i264k1.0.gYM5bQ4NgHQ)

### Top & Bottom Countries by Growth Class
```{r warning=FALSE}
# classify countries into 3 groups
gdp_capita_5yr$class <- ifelse(gdp_capita_5yr$X2006 > 20000, "Upper", "Middle")
gdp_capita_5yr$class <- ifelse(gdp_capita_5yr$X2006 < 2000, "Under", 
                               gdp_capita_5yr$class)

# factorize and put the classes in order
gdp_capita_5yr$class <- as.factor(gdp_capita_5yr$class)
gdp_capita_5yr$class <- factor(gdp_capita_5yr$class, 
                               levels(gdp_capita_5yr$class)[c(3, 1, 2)])

gdp_capita_5yr_class <- ggplot(aes(x = growth_rate, y = ..count../sum(..count..)),
                               data = gdp_capita_5yr) +
  geom_freqpoly(aes(color = class), binwidth = 5) +
  xlab("5 Years Growth Rate") + ylab("Frequency")

gdp_capita_5yr_class + scale_color_discrete(name = "")
  scale_x_continuous(limits = c(-30, 70), breaks = seq(-30, 70, 5))
```

The GDP/capita can be classified as 'Upper', 'Middle', 'Under'. With just 3 simple line plots, it is now clear that the 5 years were the time when the developed countries suffered while undeveloped countries had chance to grow. This can be yet again confirmed with the box plot that undeveloped countries have grown in the 5 years the most.

```{r}
ggplot(aes(x = class ,y = growth_rate), data = gdp_capita_5yr) +
  geom_boxplot()
```

### Conclusion

The fact found through the histogram comparison is that there are still so many countries remaining under GDP/capita 2,000 USD group while the number of countries in other groups becomes smaller as the GDP/capita increases. But it is also true that most of the countries grew over a long time period. In the short term, some undeveloped countries could grow remarkably while most developed countries struggled during the recent years as seen in the 5 years growth analysis.